% !TEX program = XeLaTeX

%\documentclass{report}
\documentclass{msu-thesis}

\usepackage[sorting=none]{biblatex}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amsfonts}
%\usepackage{url}
\usepackage{unicode-math}%, neccesary for vector graphic lowercase mathbb e, needs to use XeTex
\setmathfont[StylisticSet=3]{NewCMMath-Book.otf}

%add siunitx package for scientific notation, set spacing to tight so powers of 10 fit well
\usepackage{siunitx}
\sisetup{tight-spacing=true} 

%adds bookmarks and external linking hopefully
\usepackage[bookmarksdepth=2]{hyperref}
%set the url font to be the same as standard for MSU requirements
\urlstyle{same}

%add subfig package for side-by-side figures
%\usepackage{subfig}

\addbibresource{z_full.bib}

\begin{document}

\frontmatter

\title{Experimental Studies of Nonlinear Integrable Optics with Elliptic Potentials}
\author{John Wieland}
\date{2026}
\fieldofstudy{Physics}

\maketitlepage

\pdfbookmark{Abstract}{abstract}
\abstract{The stable transport of charged particle beams is the core challenge in designing and operating accelerators. The current paradigm for transverse focusing is based on quadrupole and dipole elements, described as a linear integrable Hamiltonian by Courant and Snyder. The operational limits of high intensity accelerators are often determined by collective instabilities within the beam. Nonlinear elements may be added to suppress certain forms of these collective effects, but restrict the range of stable trajectories. These shortcomings motivate extending to a nonlinear integrable system which can offer the benefits of suppressing collective instabilities without limiting the stable trajectories. Danilov and Nagaitsev proposed a novel nonlinear integrable system with elliptical potentials, which could be implemented as magnetic elements in an accelerator. \\ Such a system has been implemented for practical verification in the integrable optics test accelerator (IOTA), a small storage ring constructed for beam dynamics studies. Electron beam studies in IOTA treat the low-emittance beam as a macroparticle for detailed probing of the expected single particle dynamics. This dissertation details new measurements of the nonlinear integrable system relevant for practical implementation. Turn-by-turn measurements of the kicked beam responses are used for phase space reconstruction and analysis of the predicted nonlinear dynamics. The stability and aperture for various nonlinear configurations are measured with beam losses. Amplitude dependent detuning, a core figure of merit for suppressing instabilities, is measured and compared with high fidelity particle tracking simulations. The synchrotron radiation images of circulating beam allow direct measurements of the topology and lifetime in configurations where nonlinear focusing terms dominate.}

\makecopyrightpage

\chapter*{Acknowledgements}
First, I would like to acknowledge my advisor Peter Ostroumov, my guidance committee and the physics department at MSU. I have benefitted greatly from the ASET program, and extensive support during unpleasant pandemic-time education neccesities. 

Thanks to all of my friends and peers at MSU and Fermilab forming a close community of good people. Many thaks to my family, near and far, for constant support through many moves. I want to acknowledge the excellent working group at FAST/IOTA for their expert support and operation of experiments and welcoming environment.

Most importantly, I would like to acknowledge my Fermilab advisor Sasha Romanov. He has been a constant advocate and a wealth of knowledge on all things accelerator. I'm greatful for the long nights in the control room and in the cave working on practical problems. 

\clearpage

\pdfbookmark{Table of Contents}{toc}
\tableofcontents

\mainmatter

\include{chapter_1.tex}
\include{chapter_2.tex}
\include{chapter_5.tex}
\include{chapter_3.tex}
\include{chapter_4.tex}
\include{chapter_6.tex}

\begingroup
\raggedright
\printbibliography
\endgroup

\begin{appendices}
%hard coded additions to single space the appendix TOC entries
\addtocontents{toc}{\protect\vskip1.0em}
\addtocontents{toc}{\setlength{\cftbeforechapterskip}{0.0em}}
\include{appendix_1.tex}
\include{appendix_2.tex}
\include{appendix_3.tex}
\include{appendix_4.tex}
\end{appendices}

\end{document}
